Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-09-25 writeups/dissertation/dissertation.tex
foo

\section{Introduction}

@@

\section{Theoretical Background}

@@

\section{Models}

@@

\section{Method}

@@

\section{Past User Behavior}

@@

\section{Combining Predictors}

@@

@@

\section{Results for Combining Predictors}

@@

@@

\subsection{General Results}

Table \ref{tabOverallResults} summarizes model accuracies for full models (all components) on all of the datasets tested in the study.

@@

After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
Both models performed nearly equal for the StackOverflow datasets tested,
and both models were much more accurate when tested on StackOverflow compared to Twitter.

\section{General Discussion}

@@

>>>>

\section{Chapter 1: Introduction}

@@

\subsection{Roadmap}

This document is separated into 7 chapters.
Chapter 1 introduced the declarative memory research and motivated why it is interesting both from a theoretical and applied perspective to evaluate declarative memory models on large-scale datasets.
The task to test the models was also introduced: tagging of user-generated content on StackOverflow and Twitter.
Chapter 2 describes the vector-based and Bayesian declarative memory models that will be evaluated in this research, and shows how these models are related to the other common declarative memory models that are studied.
Chapter 3 presents the overall method for testing the models on the StackOverflow and Twitter datasets.
Chapter 4 shows the results for the first test of the models, where the model term for past user behavior is examined in isolation, and tags are predicted based purely on a user's past tagging history.
Chapter 5 adds the context component to the models, and shows how model accuracy is influenced by changing various architectural components of each model.
Chapter 6 describes model results when prior and context are combined and tested on both of the dataset subsets used for Chapter 4 and Chapter 5.
Finally, Chapter 7 provides a general discussion of the overall findings.

\clearpage
\section{Chapter 2: Theoretical Background}

The ACT-R declarative memory theory and the vector-based retrieval models will be described.
The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will be included.
The theory of the ACT-R and vector-based models will be compared.
Also, different tagging models have already been applied to various recommendation systems to tag user-generated content,
so these systems will be described and the models used will be related back to the Bayesian and vector-based declarative memory models.

@@

\clearpage
\section{Chapter 3: Overall Method}

Each specific result of the research (to be presented in Chapters 4-6) requires a sightly different method to study.
However, there is a common general methodology that applies across all results that will be described first.
Different subsets of the StackOverflow and Twitter datasets were used evaluate past user behavior in isolation and to test the context component of each memory model.
These datasets will be named and described in general terms.
Also, the general methods used to tokenize and stem the words in all of the datasets will be described.

\subsection{Models}

@@

@@

\clearpage
\section{Chapter 4: Past User Behavior}

@@

\clearpage
\section{Chapter 5: Combining Predictors}

@@

Chapter 5 will examine how model accuracy is influenced by two methodological concerns (adding an offset term and examining the effect of dataset subset for Twitter)
and five architectural concerns
(method of handling stop words, influence of dataset size, amount of compression used for the vector-based models, method to combine the prior and context for the vector-based models, and word order).
The two methodological concerns were examined to ensure that the technique to calibrate model parameters was accurate and also that results for Twitter were not specific to the subset of data collected.

The five architectural concerns were examined for the following reasons:
During exploratory evaluation, model accuracy seemed to be highly sensitive to the method of dealing with the commonly-used stop words, so several different methods were examined in more detail.
Due to research from \textcite{Budiu2007}, model accuracy for both the Bayesian and LSA models is influenced by the number of documents in the dataset,
so the effect of dataset size on accuracy was examined in more detail.
Specifically,
since both the Bayesian and random-permutation models are being evaluated, analyzing the effect of dataset size on accuracy in this research will allow direct comparisons between the two models on a common dataset.
This has not yet been possible since the two models have never been compared on the same dataset within the same study.
\textcite{Sahlgren2008} showed how different levels of compression in the random-permutation model influence accuracy, so this was examined for three reasons:
to replicate the findings,
as an additional check to ensure that the random-permutation model was implemented correctly, and also to see if the same levels of compression are valid for the much larger datasets used in this research.
One of the primary goals of this research is to identify the best method to combine the model term for past user behavior with the context term for the random-permutation model,
so two different ways to add the model terms were tested.
Finally, previous research has shown that including word order in the vector-based models improves performance \parencites{Jones2007, Sahlgren2008},
so word order was also examined to see if the effect could be replicated and also to measure the overall importance of word order on model accuracy relative to other architectural manipulations.

\subsection{Overall Method}

@@

@@

\clearpage
\section{Chapter 6: Combining Predictors on the Popular-Users Dataset}

@@

\clearpage
\section{Chapter 7: General Discussion}

\subsection{Overall Results}

Table \ref{tabOverallResults} summarizes model accuracies for full models (all components) on all of the datasets tested in the study.
After modifying the models by adding an entropy weighting mechanism to both models and a prior term to the random permutation model,
accuracy for the compressed random permutation model was nearly equal to the Bayesian model.
The largest difference between the two models was on the Twitter popular-hashtags dataset (7 percent).
Both models performed nearly equal for the StackOverflow datasets tested,
and both models were much more accurate when tested on StackOverflow compared to Twitter.

@@

@@

\subsection{Implications and Future Directions}
